---
title: "p8105_hw5_map2365"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
*function that checks for duplicate birthdays*
```{r}
simulate_birthdays <- function(group_size) {
birthdays <- sample(1:365, group_size, replace = TRUE)
any(duplicated(birthdays))
}
```
*running the simulation for group sizes between 2 and 50. Note, I set the seed for reproducibility purposes.*
```{r}
set.seed(1123)
group_sizes <- 2:50
n_simulations <- 10000

probabilities <- numeric(length(group_sizes))

for (i in seq_along(group_sizes)) {
  group_size <- group_sizes[i]
  results <- replicate(n_simulations, simulate_birthdays(group_size))
  probabilities[i] <- mean(results)
}
```

*ploting the results*
```{r}
plot(group_sizes, probabilities, type = "b", 
     xlab = "Group Size (n)", 
     ylab = "Probability of Shared Birthday",
     main = "Probability of At Least Two People Sharing a Birthday",
     pch = 19, col = "blue")
```
*based on the plot, we can see that the probability of two people sharing a birthday increases as the size of the group increases. Infact, when the group size reaches 50, there is nearly the probability of 2 people sharing a birthday is nearly 100%*

